### 2015-03-24 Paper1Editted/paper1Beditted.tex
updated derivation

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
Our actions often do not match our intentions when there are external disturbances such as turbulence. We derived a novel modeling approach for determining this motor intent from targeted reaching motions that are disturbed by an unexpected force. First, we demonstrated how to mathematically invert both feedforward (predictive) and feedback controls to obtain an intended trajectory. We next examined the model's sensitivity to a realistic range of parameter uncertainties, and found that the expected inaccuracy due to all possible parameter mis-estimations was less than typical movement-to-movement variations seen when humans reach to similar targets. The largest sensitivity arose mainly from uncertainty in joint stiffnesses. Humans cannot change their intent until they acquire sensory feedback, therefore we tested the hypothesis that a straight-line intent should be evident for at least the first 120 ms following the onset of a disturbance. As expected, the intended trajectory showed no change from undisturbed reaching for more than 150 ms after the disturbance onset. Beyond this point, however, we sometimes detected a change in intent, surprisingly even when the hand is already near the target. Knowing such an intent signal is broadly applicable: enhanced human-machine interaction, the study of impaired intent in neural disorders, the real-time determination (and manipulation) of error in training, and complex systems that embody planning such as brain machine interfaces, team sports, crowds, or swarms. In addition, observing intent as it changes might act as a window into the mechanisms of planning, correction, and learning.

@@

Attempts to deduce motor intent in the past have focused on the assumed spring-like properties of human muscles. Springs produce a force according to their impedance and stretch. By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point. Supporters of the ``$\lambda$ model''\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement even though it did not compensate for the dynamics of the arm as a feedforward controller would. Upon Gomi and Kawato's inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable. Therefore, it could not well-represent the intent of a simple reaching movement. Unlike the equilibrium point of muscle, the equilibrium point of the entire arm matches undisturbed movement. In the presence of disturbance, the arm might be deflected from its equilibrium. It is presently unknown if or how the arm's equilibrium point might change due to disturbances.

@@

It is important to distinguish our intended trajectory, $q_d$, from another variable often considered, the equilibrium point of the muscle actuators. Importantly, this has been claimed to be the intended movement\cite{feldman1995origin}. To consider this, we also generalize our approach to any system with a moving intention that governs the control law. This introduces two additional layers of complexity by separating the plant (arm) into its process (dynamics) and actuator components (muscles) and considering the actuator's equilibrium, $q_e$ (sometimes represented as $\lambda$). The feedforward component must use and share this actuator and hence use the actuator's equilibrium. Without loss of generality, we represent the process $P$ and actuator $A$ as linear (or linearizable) operations in some action space, $x$,
\begin{equation} \label{eq:generalPlant}
\underbrace{\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+\overbrace{\sum_{m=1}^M A_m(x-x_e)^{(m)}}^\text{Actuator}}_\text{Plant}+E=0
\end{equation}
where $x_e$ is the equilibrium trajectory of the actuator. The internal model of the process ($\hat{P}$ and $\hat{E}$) predicts system actions, external disturbances, and impedance responses in order to determine $x_e$ such that $x$ will track a desired path $x_d$.    
\begin{equation}
\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m(x_d-x_e)^{(m)}}^\text{Actuator}=0
\end{equation}
We solve this for $x_e$ and substituting into \eqref{eq:generalPlant} gives
\begin{equation}
\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+E=\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m (x_d-x)^{(m)}}^\text{Feedback (Actuator)}
\end{equation}
wherein $x_e$ vanishes, recovering our familiar model. While a proper choice of $x_e$ is perhaps essential for control, our approach does not require a model relating equilibrium and intended trajectory. Note also that the actuator's equilibrium $x_e$ is not the process's equilibrium unless all derivatives of $x_d$ are zero. This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}. In summary, if the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted and the impedances can be modeled, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.

While we treat both the process and actuator as dynamic models in our derivation, it is important to note that both the derivation and the technique are agnostic to the process model. Even if the process model was a reverberating loop \cite{fortney2012computational} driven by a reference intent, we might still recover that reference intent by simulating the loop. We are not limited to dynamic models as this might be extended to a lookup table. For instance, a switch is well-modeled as a lookup table or threshold without any consideration of its underlying mechanism. In this case, our technique could be used to determine a person's intention to flip the switch and how it changes in the face of disturbance. As long as some bidirectional relationship  exists between state and outcome (even if determined empirically), intended outcome can be determined. This allows determination of intent in many situations of interest, even where the process is irreducibly complex.

@@

\subsection*{Dynamic simulation of arm and intended trajectories}
While the derivation of intent determination is general, testing its application to human reaching requires choosing plant and actuator models, which themselves require physical parameters. Appropriate plant modeling is well-understood: measured arm segment lengths and self-reported body mass are converted into the plant's inertial, centripetal, and Coriolis terms. Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} relate body mass to limb mass, limb length to limb center of mass, and limb mass and length to moment of inertia. For the actuator, we choose the model and parameters of Burdet et al. \cite{burdet2006stability}. We fit a subject-specific constant scale factor for impedance, $c$, relative to this model to account for any task-dependence of the subjects' impedance\cite{gomi1998task}.
\begin{equation}
E=c(K_p\Delta q+K_d \Delta \dot{q})+I(q)\Delta\ddot{q}
\end{equation}
This uses linear regression, as in Franklin et al.\cite{franklin2003adaptation}, except that task-perpendicular torque and displacement from the first 150 ms of movement from noise-disturbed trials is used in place of a stationary task.

@@

\subsection*{Statistical Analysis}
Our principal goal was to determine if and when disturbed trajectories departed from undisturbed trajectories. To detect this departure, we use the one-tailed Student's t-test $(\alpha=.05)$ at 5 millisecond intervals to determine whether or not the deviation of the disturbed trajectories had exceeded the maximum deviations of the undisturbed trajectories. The span between departure of the hand and estimated intent was compared against 120 milliseconds using the sign test. The MATLAB statistics toolbox package \cite{MATLAB:2008} was used for all comparisons.

@@

\subsection*{Human Intent Trajectories}
Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (Figure 3A). If intent can change following disturbance, it can only do so after some period of time due to processing and communication delays in the sensorimotor system that produces new motor commands. We hypothesized that intent should not diverge from the undisturbed intent path even if the actual hand was disturbed within this window of delay. In both pulse timing conditions, the intended trajectory remained straight towards the target, even though the actual hand was dramatically deflected, supporting our hypothesis and the approach.

These results suggested a reproducible approach for inspecting how individuals might alter their intent in response to disturbances. Actual hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement within 145 milliseconds after the onset of disturbance. Despite the disturbance, three subjects' intents did not depart significantly from baseline. The remaining five subjects' intents deviated between 150 and 255 milliseconds after the hand deviated (Figure 3B). When disturbances occurred early in reaching, intent showed some signs of correction mid-reach. When disturbances occurred later, the intent continued to the target, but interestingly, it then corrected (Figure 3A). Intention showed systematic deflections that counteracted the direction of disturbance beginning about 200 milliseconds after the onset of the force disturbance; however, this counteraction did not immediately depart from the range of undisturbed movements (Figure 3B, orange region). This counteraction is consistent with both change of intent and stiffening in response to the disturbance.

@@

While the derivation of intent determination required many steps, the outcome is a filter whose use is straight-forward. Many filters for converting assumed intent into simulated hand trajectories have been proposed and validated \cite{shadmehr1994adaptive, franklin2003adaptation, burdet2006stability}. Demonstrating the validity and plausibility of mathematically transforming those filters to instead convert measured hand trajectories to intent adds a novel method to the arsenal of analytical methods for exploring human motor control.

The proposed approach could explain the discovery by Bizzi et al. \cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$, the equilibrium point of muscles -- $x_d$ and $\lambda$ are equivalent only when $x_d$ is unchanging; therefore, this muscle equilibrium cannot be the intent of a movement. By deriving the technique in a general form, we discovered that our $x_d$ is equivalent to $\alpha$ and can be used to determine the muscle equilibrium of a $\lambda$ model.

Interestingly, the formulation leads to some implications on the expected behavior of this moving equilibration. Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of the mass term in the intention formulation -- reaching intent cannot jump abruptly. The intent trajectories certainly follow this pattern, smoothly responding to disturbances at a latency of about 200 milliseconds. While this work examines the tendencies of the onset of change, statistical testing cannot be used to detect the absence of change in an individual movement. Still, Figure 3 shows reaches that are more compatible with intermittent control than with continuous optimal feedback control. In classic optimal control, the system is always updating intent in response to error \cite{todorov2002optimal}. While most of our subjects displayed a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}, three subjects showed no detectable response. The absence of response to a large force disturbance seems inconsistent with optimal feedback control framework and is instead consistent with adherence to a trajectory planned before the disturbance. Control that is intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent can explain both change and lack thereof. It remains to be seen if and how change in intent might be triggered.

@@

The process model does not need to be explicitly formulated. Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from swarm prediction to athletic performance. The plant (such as the actions of a crowd of people or the motions of an opposing team) may be too complex to model with simple linear transformations as we present above. However, by substituting an approximation, reverberating loop, or lookup table of input-outcome tendencies, it is possible in a variety of applications to obtain the intent, even when disturbed. In other words, it may be possible to intervene before a car crashes or an opposing team scores by knowing the intent behind their action.

@@

While we use the terms \textit{intent}, $x_d$, and \textit{equilibrium trajectory of the arm} interchangeably, they are not always identical. We assume that action does not deviate from intent in the absence of external disturbance. For this to be true, the equilibrium trajectory of the arm must be the intent. Not all action need arise from a continuous intent trajectory, however. One might volitionally twitch a muscle without planning its effect on the limb. In the context of reaching, our results and others' \cite{flash1991arm, bizzi1984posture, shadmehr1994adaptive, rohrer2004submovements, fortney2012computational, mistry2013optimal} suggest a continuous pre-planned trajectory with kinematic constraints, perhaps composed in an intermittent fashion from some motor primitive. This planned trajectory may not itself be intended and might arise from an intent to minimize cost or effort.

@@

\begin{figure}
\caption{
{\bf Subjects' disturbed hand trajectories and extraction of intended trajectories from them reveal that both variance and invariance of the motor plan can occur even in response to very large disturbances.} The hand path, in black, deviates from the blue baseline, as a force pulse, gray arrows, is applied. This subject's extracted desired trajectories, in orange, do not significantly differ from baseline during force application, but show some reaction afterwards.
}
\label{fig:anecdotes}
\end{figure}

\begin{figure}
\caption{
{\bf Comparison of subjects' perpendicular deviations across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching for at least 300 ms.} At most 135 milliseconds after the onset of disturbance, hand deviation in the presence of disturbance (black dots, black band is mean $\pm$ 95\% confidence interval on the mean) can no longer be explained by undisturbed hand deviation (blue band). The specific time at which this occurs is marked with a green line for each subject. Movement intention (orange dots, orange band is mean $\pm$ 95\% confidence interval on the mean) does not significantly deviate until at least 300 ms after the onset of disturbance. The specific time at which this occurs is marked with a red line for each subject.
}
\label{fig:grouptrends}
\end{figure}
\pagebreak

>>>>

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
Our actions often do not match our intentions when there are external disturbances such as turbulence. We derived a novel modeling approach for determining this motor intent from targeted reaching motions that are disturbed by an unexpected force. First, we demonstrated how to mathematically invert both feedforward (predictive) and feedback controls to obtain an intended trajectory. We next examined the model's sensitivity to a realistic range of parameter uncertainties, and found that the expected inaccuracy due to all possible parameter mis-estimations was less than typical movement-to-movement variations seen when humans reach to similar targets. The largest sensitivity arose mainly from uncertainty in joint stiffnesses. Humans cannot change their intent until they acquire sensory feedback, therefore we tested the hypothesis that a straight-line intent should be evident for at least the first 120 milliseconds following the onset of a disturbance. As expected, the intended trajectory showed no change from undisturbed reaching for more than 150 milliseconds after the disturbance onset. Beyond this point, however, we detected a change in intent in five out of eight subjects, surprisingly even when the hand is already near the target. Knowing such an intent signal is broadly applicable: enhanced human-machine interaction, the study of impaired intent in neural disorders, the real-time determination (and manipulation) of error in training, and complex systems that embody planning such as brain machine interfaces, team sports, crowds, or swarms. In addition, observing intent as it changes might act as a window into the mechanisms of planning, correction, and learning.

@@

Attempts to deduce motor intent in the past have focused on the assumed spring-like properties of human muscles. Springs produce a force according to their impedance and stretch. By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point. Supporters of the ``$\lambda$ model''\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement even though it did not compensate for the dynamics of the arm as a feedforward controller would. Upon Gomi and Kawato's inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable. Therefore, it could not well-represent the intent of a simple reaching movement. Unlike the equilibrium point of muscle, the equilibrium point of the entire arm is the path the arm will follow in the absence of disturbance. In the presence of disturbance, the arm might be deflected from its equilibrium. It is presently unknown if or how this arm equilibrium point might change due to disturbances.

@@

To explore the conditions under which this procedure for recovering the intent trajectory can be generalized, we must take a series of counter-intuitive steps. First, we distinguish between joint space and generalized space by introducing a generalized state variable, $x$. Second, we separate the plant (arm) into its process (dynamics), $P$, and actuator (muscles), $A$, components. Third, we use power series expansions to model these components in a very general sense,
\begin{equation} \label{eq:generalPlant}
\underbrace{\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+\overbrace{\sum_{m=1}^M A_m(x-x_e)^{(m)}}^\text{Actuator}}_\text{Plant}+E=0
\end{equation}
where $x_e$ is the equilibrium trajectory of the actuator. Fourth, we describe the actuator control law using the same expansions,
\begin{equation}
\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m(x_d-x_e)^{(m)}}^\text{Actuator}=0
\end{equation}
where an internal model, $\hat{P}$ and $\hat{E}$, predicts system dynamics and external disturbances in order to determine the actuator equilibrium $x_e$ such that $x$ will track a desired path $x_d$. The feedforward component must use the actuator and hence it shares the actuator's equilibrium with the plant. This allows us to finally relate the physical system and its control law by solving for $x_e$ and substituting into \eqref{eq:generalPlant},
\begin{equation}
\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+E=\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m (x_d-x)^{(m)}}^\text{Feedback (Actuator)}
\end{equation}
wherein $x_e$ vanishes recovering our familiar model. While a proper choice of $x_e$ is perhaps essential for control, our approach does not require a model relating this equilibrium and intended trajectory. Note also that the actuator's equilibrium $x_e$ is not the process's equilibrium unless all derivatives of $x_d$ are zero. This is important as the actuator state ($\lambda$) has been claimed to be the intended movement\cite{feldman1995origin}. This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}. If the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted and the impedances can be modeled, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.

While we treat both the process and actuator as dynamic models in our derivation, it is important to note that both the derivation and the technique are agnostic to the process model. Instead of a dynamic model, it could be a lookup table. For instance, a switch is well-modeled as a lookup table or threshold without any consideration of its underlying mechanism. In this case, our technique could be used to determine a person's intention to flip the switch and how it changes in the face of disturbance. As long as some bidirectional relationship exists between state and outcome (even if determined empirically), intended outcome can be determined. This allows determination of intent in many situations of interest, even where the process is irreducibly complex.

@@

\subsection*{Dynamic simulation of arm and intended trajectories}
While the derivation of intent determination is general, testing its application to human reaching requires choosing plant and actuator models, which themselves require physical parameters. Appropriate plant modeling is well-understood: measured arm segment lengths and self-reported body mass are converted into the plant's inertial, centripetal, and Coriolis terms. Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} relate body mass to limb mass, limb length to limb center of mass, and limb mass and length to moment of inertia. For the actuator, we choose the model and parameters of Burdet et al. \cite{burdet2006stability}. We fit a subject-specific constant scale factor for impedance, $c$, relative to this model to account for any task-dependence of the subjects' impedance\cite{gomi1998task}.
\begin{equation}
c\big(K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})\big)=\big(\hat{M}(q_d)\ddot{q}_d+\hat{G}(q_d,\dot{q}_d)\big)-\big(M(q)\ddot{q}+G(q,\dot{q})\big)+E
\end{equation}
This uses linear regression, as in Franklin et al.\cite{franklin2003adaptation}, except that task-perpendicular torque and displacement from the first 150 ms of movement from noise-disturbed trials is used in place of a stationary task.

@@

\subsection*{Statistical Analysis}
Our principal goal was to determine if and when disturbed trajectories departed from undisturbed trajectories. To detect this departure, we use the one-tailed Student's t-test $(\alpha=.05)$ at 5 millisecond intervals to determine whether or not the deviation of the disturbed trajectories had exceeded the maximum deviations of the undisturbed trajectories. The time span between departure of the hand and estimated intent was compared against 120 milliseconds using the nonparametric \textit{sign test}. The MATLAB statistics toolbox package \cite{MATLAB:2008} was used for all comparisons.

@@

\subsection*{Human Intent Trajectories}
Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (Figure 3). If intent can change following disturbance, it can only do so after some period of time due to processing and communication delays in the sensorimotor system that produces new motor commands. We hypothesized that intent should not diverge from the undisturbed intent path even if the actual hand was disturbed within this window of delay. In both pulse timing conditions, the intended trajectory remained straight for longer than the hypothesized 120 milliseconds ($p<0.001$), even though the actual hand was dramatically deflected, supporting our hypothesis and the approach.

These results suggested a reproducible approach for inspecting how individuals might alter their intent in response to disturbances. Actual hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement within 145 milliseconds after the onset of disturbance. Despite the disturbance, three of the eight subjects' intents did not depart significantly from baseline (Figure 3A). The remaining five subjects' intents deviated between 150 and 255 milliseconds after the hand deviated (Figure 3B). When disturbances occurred early in reaching, intent showed some signs of correction mid-reach. When disturbances occurred later, the intent continued to the target, but interestingly, it then corrected. Intention showed systematic deflections that counteracted the direction of disturbance beginning about 200 milliseconds after the onset of the force disturbance; however, this counteraction did not immediately depart from the range of undisturbed movements (Figure 3B, bottom panels).

@@

While the derivation of intent determination required many steps, the outcome is a filter whose use is straight-forward. Many filters for converting assumed intent into simulated hand trajectories have been proposed and validated \cite{shadmehr1994adaptive, franklin2003adaptation, burdet2006stability}. Demonstrating the validity and plausibility of mathematically transforming those filters to instead convert measured hand trajectories to intent adds a novel method to the arsenal of analytic methods for exploring human motor control.

The proposed approach could explain the discovery by Bizzi et al. \cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$, the equilibrium point of muscles -- $x_d$ and $\lambda$ are equivalent only when $x_d$ is unchanging; therefore, this muscle equilibrium cannot be the intent of a movement. By deriving the technique in a general form, we discovered that our $x_d$ is equivalent to the virtual trajectory and can be used to determine the muscle equilibrium of a $\lambda$ model.

Interestingly, the formulation leads to some implications on the expected behavior of this moving equilibrium. Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of the mass term in the intention formulation -- reaching intent cannot jump abruptly. The intent trajectories certainly follow this pattern, smoothly responding to disturbances at a latency of about 200 milliseconds. While this work examines the tendencies of the onset of change, statistical testing cannot be used to detect the absence of change in an individual movement. Still, Figure 3 shows reaches that are more compatible with intermittent control than with continuous optimal feedback control. In classic optimal control, the system is always updating intent in response to error \cite{todorov2002optimal}. While most of our subjects displayed a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}, three of our eight subjects showed no detectable response. The absence of response to a large force disturbance is instead consistent with adherence to a trajectory planned (perhaps entirely) before the disturbance. Control that is intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent can explain such observations. It remains to be seen if and how change in intent might be triggered.

@@

The process model does not need to be explicitly formulated. Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from swarm prediction to athletic performance. The plant (such as the actions of a crowd of people or the motions of an opposing team) may be too complex to model with simple linear transformations as we present above. However, by substituting any approximation or lookup table of input-outcome tendencies, it is possible in a variety of applications to obtain the intent, even when disturbed. In other words, it may be possible to intervene before a car crashes or an opposing team scores by knowing the intent behind their action.

@@

The words \textit{intent} and \textit{equilibrium} are contentious because of the various definitions that exist in the literature. Our simple definition here is that the intent is the path that would have been taken had there been no external disturbances. A rich history of research has suggested many ways that an equilibrium trajectory might explain how the central nervous system \cite{feldman1995origin} and/or pre-planned elements \cite{shadmehr1994adaptive} can explain action. Here we suggest that one should directly define intent in the context of a hypothesized control model. Our work here assumed the model of Burdet \cite{burdet2006stability}. However, other, more sophisticated, models may be deemed more appropriate in the future. What is clear is that given some model of control, this intent extraction approach is able to invert the process to estimate intent.

It is important to mention that it is not necessary for the brain to represent intent explicitly to generate action. It is possible that the nervous system can simply learn the relationship between muscle activations and the accomplishment of goals. For example, one might simply evaluate a motion outcome and adjust descending signals to the muscles. At the same time, one may not deny the fact that there are tasks that require a explicit representation of a trajectory, such as performing a dance, drawing a picture, or conducting an orchestra. For these tasks, intent is likely represented at  some level in the nervous system. Nevertheless, the mechanics of the arm dictate the existence of an equilibrium whether intent is explicitly represented or not. Here our goal was more modest, which was to answer, "where would the arm have gone?"

@@

\begin{figure}
\caption{
{\bf Subjects' disturbed hand trajectories and extraction of intended trajectories from them reveal that both variance and invariance of the motor plan can occur even in response to very large disturbances.} The hand path, in black, deviates from the blue baseline, as a force pulse, gray arrows, is applied. Three of the eight subjects' intent trajectories (orange) did not significantly deviate from undisturbed movement (panel A) as observed by the intent (orange) region (mean $\pm$ 95\% confidence interval on the mean) not departing the undisturbed movement (blue) region. The time at which the hand (black) region departed from undisturbed movement is marked with a green dots. The time of hand departure for the two subjects with similar trajectories is also marked with green dots. Five of the eight subjects' intent trajectories did significantly deviate from undisturbed movement (panel B). The time of intent departure from the undisturbed trajectories is marked with red dots.
}
\label{fig:anecdotes}
\end{figure}

\pagebreak
